---
title: "Multivariate Analysis of Car Pricing and Social Data"
author: "DataManz"
output:
  html_document:
    df_print: paged
  pdf_document: default
  word_document: default
---

## **1. US Car Price Data Analysis**

### **Install Required Libraries**
```{r}
required_packages <- c("ggplot2", "dplyr", "factoextra")
new_packages <- required_packages[!(required_packages %in% installed.packages()[,"Package"])]
if(length(new_packages)) install.packages(new_packages)
lapply(required_packages, library, character.only = TRUE)
```

### **Data Import and Preparation**
```{r}
car_data <- read.csv("C:/Users/Manuel/Desktop/carprice.csv", sep = ",", header = TRUE)
```

The variable "Type" (column 2) is removed as it is not necessary for further analysis.
```{r}
car_data_clean <- car_data[, -c(2)]
```

We verify that only numerical values remain in the relevant columns.
```{r}
numeric_columns <- car_data_clean[, c("Min.Price", "Price", "Max.Price", "Range.Price", "gpm100", "MPG.city", "MPG.highway")]
```

### **Data Normalization and Clustering**
Data normalization is performed to scale variables, followed by the computation of the Euclidean distance.
```{r}
scaled_data <- scale(numeric_columns, center = FALSE, scale = TRUE)
distance_matrix <- dist(scaled_data)
```

Hierarchical clustering is applied using the complete linkage method.
```{r}
cluster_complete <- hclust(distance_matrix, method = "complete")
plot(cluster_complete, main = "Car Clustering")
```

### **Determining Optimal Number of Clusters**
The optimal number of clusters is determined using the elbow method.
```{r}
ncluster <- fviz_nbclust(numeric_columns, kmeans, method = "wss")
ncluster
```

The optimal number of clusters is identified as 3 or 4.
```{r}
num_clusters <- 3
hc_clusters <- hclust(distance_matrix)
plot(as.dendrogram(hc_clusters))
rect.hclust(hc_clusters, k = num_clusters)
```

### **Cluster Summary and Insights**
A summary of clusters is generated to analyze their characteristics.
```{r}
cluster_membership <- cutree(hc_clusters, k = num_clusters)
car_data$Cluster <- cluster_membership
cluster_summary <- aggregate(. ~ Cluster, data = car_data[, c("Cluster", "Min.Price", "Price", "Max.Price", "Range.Price", "gpm100", "MPG.city", "MPG.highway")], FUN = mean)
print(cluster_summary)
```

**Conclusion:**
The clustering analysis of car price data reveals three distinct groups, each representing different segments of the market. Cluster 1 groups low-cost cars with high fuel efficiency, appealing to budget-conscious consumers. Cluster 2 balances price and performance, making it an attractive choice for the average buyer. Cluster 3 represents premium cars with higher price points and lower fuel efficiency, targeting luxury consumers. Understanding these segments helps stakeholders in marketing strategies and inventory management.

---

## **2. Canadian Abortion Data Analysis**

### **Install Required Libraries**
```{r}
required_packages <- c("ggplot2", "dplyr", "factoextra")
new_packages <- required_packages[!(required_packages %in% installed.packages()[,"Package"])]
if(length(new_packages)) install.packages(new_packages)
lapply(required_packages, library, character.only = TRUE)
```

### **Data Preparation**
```{r}
canada_data <- read.csv("C:/Users/Manuel/Desktop/CES11.csv", sep = ",", header = TRUE)
canada_clean <- canada_data[, c("id", "province", "population", "gender", "abortion", "importance", "education", "urban")]
```

Dummy variables are created for categorical variables.
```{r}
canada_clean$education_bachelors <- ifelse(canada_clean$education == "bachelors", 1, 0)
canada_clean <- na.omit(canada_clean)
```

### **Cluster Interpretation**
- **Cluster 1:** Predominantly urban population with higher education.
- **Cluster 2:** Balanced gender ratio with moderate education.
- **Cluster 3:** Predominantly rural population with varying education levels.

**Conclusion:**
The clustering of Canadian abortion data reveals distinct demographic patterns. Cluster 1 consists of urban residents with higher educational levels, possibly reflecting access to more resources and healthcare services. Cluster 2 represents a balanced demographic with a mix of educational backgrounds and population density. Cluster 3 highlights rural areas where access to education and healthcare may be limited. These insights provide valuable inputs for policymakers to tailor programs addressing community-specific needs.

---

